\myTitle

### 2023-11-28 body-text-of-paper.tex
MGD added TJS's changes

Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative 
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations 
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker 
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently 
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location 
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications, 
such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics 
\citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.

In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of 
mixed ancestry---
characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu 
\citet{pritchard2000inference}.  Various processes could contribute to such mixed associations.  Most directly, individuals may represent the 
influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately 
representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of 
population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described 
as discrete genetic clusters.
NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, 
something like END NOTE
Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source 
populations not sufficiently
represented in the sample to be among the characterized $K$ subpopulations.
Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data,
as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex 
 ancestry patterns were created.

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on simulated data where the truth is known, structured to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see CITE, CITE, CITE).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed 
samples, and
inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
documented this type of artifact in the context of simulations to assess the power for
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
We demonstrate, below, that similar artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
power inflation, or RIPI for short.
NOTE I would proposed holding on to the next sentence as we build the idea of mixed ancestry in the next paragraph END NOTE
In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
from simulations by constraining the genetic differences among simulated populations to match
those of populations from which the samples were obtained.

@@

In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without 
replacement, thus maintaining the genetic differences in the empirical dataset.
Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of 
varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes.

@@

Our approach for simulating admixed individuals without inflating perceived power for
the estimation of their admixture fractions is motivated by the traditional cross-validation
approach for evaluating the power of mixture proportion
estimation in genetic stock identification (e.g., the cross-validation approach
desribed in \citealt{moran2019bayesian}).
In the cross-validation approach, individuals are removed from the reference
samples and placed in a mixture sample of "unknowns", whose origin
is inferred using the remaining individuals in the reference samples.  A key
feature of this approach is that no new genetic material is being created by
sampling with replacement; however the genetic material of all individuals
in the original samples is being used---either within the remaining individuals in the
reference samples or within the individuals in the mixture.

Our scheme, described here, extends the traditional cross-validation approach
by simulating the descent of chunks of chromosome through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or to
remain within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.

We call the structure
we use to do this sort of cross-validation simulation of admixed individuals
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
following recombination between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map,
 one gamete is assorted down each edge from the founder node to its daughter.
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material found in the founders (and
no more than that)
but that material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) respects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individuals and markers available to
use in downstream analyses.

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2}); 
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $BC_1$ in sample node~s9; and two $BC_2$'s in sample node~s10. 
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\gspcaptwo}
\label{fig:gsp2}
\end{figure}
%%%%
however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
Before simulating from a GSP, \gscramble{} checks to ensure the GSP
satisfies these validity conditions.  The simulation of genomic segments
from founders to samples in a GSP is performed by the \gscramble{} function
{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
to segregate genomic segments, and, for each such GSP, a table that maps the
population labels on each GSP (in the inverted triangles atop the founders) to the
different sampled groups of individuals.

@@

These permutations are executed upon the genotypes provided by the user.
\gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
$2L$ columns (two for each locus). If the data have been phased, then the phased alleles from the
first haplotype should occupy the
first entry for each locus within an individual, while alleles phased together on the second haplotype should occupy the
second entry for each locus. 
There can be several populations represented in the data set, but
each individual must belong to exactly one
population.

@@

Permutation of alleles is handled within the \gscramble{} function {\footnotesize\tt segments2markers()}
which takes as input the segments segregated down a GSP, a matrix of sampled genotypes, and a specification
of the group that each sampled individual belongs to.  The permutation performed when running
{\footnotesize\tt segments2markers()} is controlled by two main options.  The first, {\tt\footnotesize preserve\_individuals},
can take three different values.  If {\tt\footnotesize preserve\_individuals = TRUE}, then entire individuals are permuted,
such that all the alleles within an individual remain together.  If {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, then
the alleles within an individual found on a homologous pair of chromosomes remain together, but the different
chromosome pairs of an individual are independently permuted into new positions in the data set.  This form of permutation
maintains the LD between markers on the same chromosome, which may be desirable in some contexts.
If {\tt\footnotesize preserve\_individuals = FALSE} then each gene copy within an individual is permuted separately,
providing the maximal scrambling of the data.   The second option, {\tt\footnotesize preserve\_haplotypes}, can take two
different values, {{\tt\footnotesize FALSE} (the default) and {\tt\footnotesize TRUE}.  This option, which keeps alleles together
on the same haplotype, should be set to true only
if the original genotypes are phased into separate haplotypes.

@@

Genetic samples were collected from 160 invasive wild pigs in 
southeastern Missouri that were lethally removed through population control efforts conducted 
as a component of the Missouri Feral Hog Elimination Partnership
by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation, 
and other cooperative state agencies CHECK.
DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with 
commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).
Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides 
62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes.
We implemented standard genotype quality control metrics, specifically, we pruned genotypes with 
call rates $<95\%$ and minor allele frequencies $<5\%$.
We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size 
of 5 loci to remove markers above a linkage threshold of $R^2 > 0.5$.

The 160 invasive wild pig samples represented 120 reference individuals from the core regions of 
3 populations (40 individuals each from Pop1, Pop2, Pop14) and 40 individuals from the contact regions of these populations. 
We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure of the 160 samples (for $K = {1...6}$)
and identify that indeed the 120 reference individuals represented 3 distinct genetic clusters.

Using the 120 reference individuals, we sought to classify the admixture patterns of the 40 empirical samples 
from the contact regions using simulated pedigrees from \gscramble{}. To address this objective we simulated 
pedigrees to represent the likely classes of admixed individuals that would be present in the samples 
from the contact regions of the 3 reference populations. The types of admixed individuals we considered were 
all two-population combinations of $F_1$, $F_2$, $BC_1$ and $BC_2$. We also 
simulated individuals from each of the founder populations. The first step of this workflow involved 
simulating admixed individuals for 2 founder populations using one of the preset \gscramble{} 
GSPs provided by the function {\footnotesize\tt create\_GSP()} for producing each of the types of admixed 
individuals. While we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}), 
different numbers of simulated individuals are produced for each hybrid class. Therefore, for the sake of the 
computational efficiency from simulating the exact number of desired individuals, we chose to simulate from 
a single GSP for each of the hybrid classes by setting the corresponding hybrid parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE}. We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()} to specify the recombination rates. The 120 simulated genotypes were written to PLINK .ped and .map files using {\footnotesize\tt gscramble2plink()}. We used PLINK version 1.9 \citep{purcell2007plink} to convert the genotypes to binary files and used these as input into ADMIXTURE for which we specified $K=3$ to produce ancestry values. We iteratively 
performed this workflow to produce 1,000 simulated genotypes and ancestry values for each of the 
4 hybrid classes, as well as founder individuals, for each of the population combinations, for a total of 
21,000 simulated individuals.

\gscramble{} REFERENCE GROUPS
\gscramble{} PEDIGREES
ADMIXTURE \gscramble{} VS EMPIRICAL
FST OF EMPIRICAL AND \gscramble{} CORE POPULATIONS

@@

\section*{Acknowledgements}
We thank etc. etc.   Contribution number  mHAVEAGAS-003. 
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder. 
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University, 
and the National Science Foundation (award 2201538).

>>>>

Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications,
such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics and affecting patterns 
of invasive species expansion \citep{huestis2019windborne, estoup2010reconstructing}.

In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of
mixed ancestry--- individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}. 
Various processes could contribute to such mixed associations.  Most directly, individuals may represent the
influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately
representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of
population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described
as discrete genetic clusters.
Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source
populations not sufficiently
represented in the sample to be among the characterized $K$ subpopulations.
Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data,
as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex
 ancestry patterns were created.

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on simulated data where the truth is known, yet structured to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see CITE, CITE, CITE).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
samples, and
inflates the perceived statistical power to correctly assign individuals back to source populations. 
We refer to this phenomenon as resampling-induced power inflation (RIPI).
\citet{anderson2008improved} used simulations to document this type of artifact in the context of 
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
Similar statistical artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.

@@

In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
replacement, thus maintaining the genetic differences in the empirical dataset and eliminating the effect of RIPI.
Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of
varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes.

@@

Our approach for simulating admixed individuals without inflating perceived power for
the estimation of their admixture fractions is motivated by the traditional cross-validation
approach for evaluating the power of mixture proportion
estimation in genetic stock identification (e.g., the cross-validation approach
described in \citealt{moran2019bayesian}).
In the cross-validation approach, individuals are removed from the reference
samples and placed in a mixture sample of "unknowns", whose origin
is inferred using the remaining individuals in the reference samples.  A key
feature of this approach is that no new genetic material is being created by
sampling with replacement; rather the genetic material of all individuals
in the original sample is being used---either within the remaining individuals in the
reference samples or within the individuals in the mixture.

Our scheme, described here, extends the traditional cross-validation approach
by simulating the descent of chromosomal linkage blocks through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or is retained 
within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.

We refer to the structure used to implement this cross-validation simulation 
of admixed individuals
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
following recombination between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map,
 one gamete is assorted down each edge from the founder node to its daughter.
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material (and no more) found among the founders but that 
material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) reflects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individuals and markers available to
use in downstream analyses.

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2});
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $BC_1$ in sample node~s9; and two $BC_2$'s in sample node~s10.
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\gspcaptwo}
\label{fig:gsp2}
\end{figure}
%%%%
however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating a unique GSP, 
the following requirements are necessary for the GSP to be valid 
and the simulation to execute successfully. 
Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
Before simulating from a GSP, \gscramble{} checks to ensure the GSP
satisfies these validity conditions.  The simulation of genomic segments
from founders to samples in a GSP is performed by the \gscramble{} function
{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
to segregate genomic segments, and, for each such GSP, a table that maps the
population labels on each GSP (in the inverted triangles atop the founders) to the
different sampled groups of individuals.

@@

These permutations are implemented with the genotypes provided by the user.
\gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
$2L$ columns (two for each locus). If the data have been phased, then the phased alleles from the
first haplotype should occupy the
first entry for each locus within an individual, while alleles phased together on the second haplotype should occupy the
second entry for each locus.
There can be several populations represented in the data set, but
each individual must belong to exactly one
population.

@@

Permutation of alleles is handled within the \gscramble{} function {\footnotesize\tt segments2markers()}
which takes as input the segments segregated down a GSP, a matrix of sampled genotypes, and a specification
of the group that each sampled individual belongs to.  The permutation performed when running
{\footnotesize\tt segments2markers()} is controlled by two main options.  The first, {\tt\footnotesize preserve\_individuals},
can take three different values.  If {\tt\footnotesize preserve\_individuals = TRUE}, then entire individuals are permuted,
such that all the alleles within an individual remain together.  If {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, then
the alleles within an individual found on a homologous pair of chromosomes remain together, but the different
chromosome pairs of an individual are independently permuted into new positions in the data set.  This form of permutation
maintains the linkage disequilibrium (LDL between markers on the same chromosome, which may be desirable in some contexts.
If {\tt\footnotesize preserve\_individuals = FALSE} then each gene copy within an individual is permuted separately,
providing the maximal scrambling of the data. The second option, {\tt\footnotesize preserve\_haplotypes}, can take two
different values, {{\tt\footnotesize FALSE} (the default) and {\tt\footnotesize TRUE}. 
This option, which keeps alleles together
on the same haplotype, should be set to true only
if the original genotypes are phased into separate haplotypes.

@@

Genetic samples were collected from 160 invasive wild pigs in
southeastern Missouri, lethally removed through population control efforts conducted
as a component of the Missouri Feral Hog Elimination Partnership
by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation,
and other cooperative state agencies.
DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with
commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).
Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides
62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes.
We implemented standard genotype quality control measures; specifically, we pruned loci with
call rates $<95\%$ and minor allele frequencies $<5\%$.
We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size
of 5 loci to remove markers above a linkage threshold of $R^2 > 0.5$.

The 160 invasive wild pig samples represented 120 reference individuals from the core regions of
3 populations (40 individuals each informed by previous analyses) 
and 40 individuals from the contact regions of these populations.
We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure 
of the 160 samples (for $K = {1...6}$). Using the 120 reference individuals identified to represent population cores, 
we sought to classify the admixture patterns of the 40 empirical samples
from the contact regions using simulated pedigrees from \gscramble{}. To address this objective we simulated
pedigrees to represent the likely classes of admixed individuals that would be present in the samples
from the contact regions of the 3 reference populations. The types of admixed individuals we considered were 
$F_1$, $F_2$, $F_{1}BC_1$ and $F_{1}BC_2$ representing all two-population combinations 
(Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2 as required 
to simulate backcrossed classes). 
We also simulated individuals from each of the founder populations. The first step of this workflow involved
simulating admixed individuals for 2 founder populations using one of the preset \gscramble{}
GSPs provided by the function {\footnotesize\tt create\_GSP()} for producing each of the types of admixed
individuals. Although we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}),
different numbers of simulated individuals are produced for each hybrid class. Therefore, for the sake of the
computational efficiency from simulating the exact number of desired individuals, we chose to simulate from
a single GSP for each of the hybrid classes by setting the corresponding hybrid parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE}. We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()} to specify the recombination rates. The 120 simulated genotypes were written to PLINK .ped and .map files using {\footnotesize\tt gscramble2plink()}. We used PLINK version 1.9 \citep{purcell2007plink} to convert the genotypes to binary files and used these as input into ADMIXTURE for which we specified $K=3$ to produce ancestry values. We iteratively
performed this workflow to produce 1,000 simulated genotypes and ancestry values for each of the
4 hybrid classes, as well as founder individuals, for each of the population combinations, for a total of
21,000 simulated individuals.

@@

\section*{Acknowledgements}
Research was supported in part by the USDA-APHIS-National Feral Swine Damage 
Management Program. We are grateful for the numerous Wildlife Services and Missouri 
Department of Conservation employees that contributed to genetic sampling of wild pigs 
within southeastern Missouri. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis, 
Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in 
Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University,
and the National Science Foundation (award 2201538).
